Multiresolution Neural Field Estimation

### 2011-03-03 Multiresolution_framework/ltx/manuscript.tex
edits to result section

To derive the state-space model, Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we cross-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\int_{\Omega} \boldsymbol\mu  \left(r\right)\boldsymbol\mu\left(r\right)^\top  dr\mathbf{x}_{t+1}= 
	\xi\int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu\left(r\right)^\top dr\mathbf{x}_t \nonumber \\
	&+T_s \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r'\right)\mathbf{x}_t 
	\, dr'dr} \nonumber \\
	+& \int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
In order to simplify the model we define the matrices
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t 
	\, dr'dr}
\end{align}
and substitute them into Eq.~\eqref{eq:DecomposedModel2}. Then we cross-multiply by $\mathbf{\Lambda}_{x}^{-1}$ giving
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t \\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= (T_s\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I),
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

@@

Field reconstruction \dean{using the results of the EM algorithm} at different levels of approximations is shown in \figurename{\ref{fig:FieldEstimates}}. \dean{Needs more info.} The figure confirms that $j=2$ is adequate to capture the dynamic of the underlying field \dean{since the error converges?}. The root mean square error (RMSE) of the field estimates is shown in \figurename{\ref{fig:ParametersHistogram}} (a).

The histogram in \figurename{\ref{fig:ParametersHistogram}} demonstrates the performance of the parameter estimation. \dean{The results are good because the RMSE between the true and estimated kernels is $1.83$ (units?)... There is a bias of 7.4 \% in $\theta_1$ because..., which is not big deal.}
% \begin{figure}[!h]
%  \centering
%  \includegraphics[scale=.9]{./Graph/ObservationFreqResponse.pdf}
%  \caption{(a) Mexican-hat connectivity kernel. The kernel is a composite of two B-spline
% functions (dashed lines), representing local excitation and lateral inhibition. (b) The average (over time) power in dB of the spatial frequency of the observations. The dashed line shows the cutoff frequency (-3 dB point).}
% \label{fig:KernelAndFreqResponse} 
%   \end{figure} 
\begin{figure}[!h] 
 \centering
 \includegraphics[scale=0.9]{./Graph/Field.pdf}
 \caption{Actual (solid line) and estimated (dashed line) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, RMSE = . (b) $j=1$, $n_x=33$, RMSE = . (c) $j=2$, $n_x=65$, RMSE = . (d) $j=3$, $n_x=131$, RMSE = . (e) $j=4$, $n_x=263$, RMSE = .}
 \label{fig:FieldEstimates}
 \end{figure} 
% \begin{figure}[!h] 
%  \centering
%  \includegraphics[scale=1]{./Graph/Hist.pdf}
%  \caption{(a) RMSE of the estimated field at different spatial scales. (b-c) Histograms of the excitatory and inhibitory connectivity basis function parameter estimates, $\theta_0$ and $\theta_1$, respectively. The solid lines show the actual parameter values and the dotted lines show the estimated parameters means.}
%  \label{fig:ParametersHistogram}
%  \end{figure} 
\begin{figure}[!h] 
 \centering
 \includegraphics[scale=1]{./Graph/KernelEstimate.pdf}
 \caption{The actual connectivity kernel is shown the the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std). The 7 weighted kernel basis functions are shown by the black dashed lines.}
 \label{fig:KernelEstimate}
 \end{figure}

>>>>

\parham{To derive the state-space model, we first define the matrices}
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t 
	\, dr'dr}
\end{align}
Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are then substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we cross-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Then we cross-multiply by $\mathbf{\Lambda}_{x}^{-1}$ giving
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= (T_s\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I),
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

@@

Field reconstruction \dean{using the results of the EM algorithm} at different levels of approximations is shown in \figurename{\ref{fig:FieldEstimates}}. \dean{Needs more info.} The figure confirms that $j=3$ is adequate to capture the dynamic of the underlying field \dean{since the root mean square error (RMSE) converges to 1.04.}

The histogram in \figurename{\ref{fig:ParametersHistogram}} demonstrates the performance of the parameter estimation. \dean{The results are good because the RMSE between the true and estimated kernels is $1.83$ (units?)... There is a bias of 7.4 \% in $\theta_1$ because..., which is not big deal.}
% \begin{figure}[!h]
%  \centering
%  \includegraphics[scale=.9]{./Graph/ObservationFreqResponse.pdf}
%  \caption{(a) Mexican-hat connectivity kernel. The kernel is a composite of two B-spline
% functions (dashed lines), representing local excitation and lateral inhibition. (b) The average (over time) power in dB of the spatial frequency of the observations. The dashed line shows the cutoff frequency (-3 dB point).}
% \label{fig:KernelAndFreqResponse} 
%   \end{figure} 
\begin{figure}[!h] 
 \centering
 \includegraphics[scale=1]{./Graph/Field.pdf}
 \caption{Actual (black line) and estimated (gray line) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, RMSE =4.44 . (b) $j=1$, $n_x=33$, RMSE =1.87 . (c) $j=2$, $n_x=65$, RMSE =1.2 . (d) $j=3$, $n_x=131$, RMSE = 1.04 . (e) $j=4$, $n_x=263$, RMSE = 1.04 .}
 \label{fig:FieldEstimates}
 \end{figure} 
% \begin{figure}[!h] 
%  \centering
%  \includegraphics[scale=1]{./Graph/Hist.pdf}
%  \caption{(a) RMSE of the estimated field at different spatial scales. (b-c) Histograms of the excitatory and inhibitory connectivity basis function parameter estimates, $\theta_0$ and $\theta_1$, respectively. The solid lines show the actual parameter values and the dotted lines show the estimated parameters means.}
%  \label{fig:ParametersHistogram}
%  \end{figure} 
\begin{figure}[!h] 
 \centering
 \includegraphics[scale=1]{./Graph/Kernel.pdf}
 \caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std). The 7 weighted kernel basis functions are shown by the black dashed lines.}
 \label{fig:KernelEstimate}
 \end{figure}
